---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
author: Eva Yifan Gong
---

## Loading and preprocessing the data
In initial processing of the data, I remove NAs and transform `date` variable into Date object.
```{r}
library(dplyr)
data=read.csv("./Data/activity.csv")
data$date <-as.Date(data$date)
data_cleaned <- data[complete.cases(data$step),] #remove NAs
```

## What is mean total number of steps taken per day?
The mean total number of steps taken per day is `10766.19`, and the median is `10765`.
```{r}
step_per_day <- data_cleaned %>%
  group_by(date) %>%
  summarize(total=sum(steps,na.rm=TRUE))
hist(step_per_day$total)
mean(step_per_day$total)
median(step_per_day$total)
```


## What is the average daily activity pattern?
Interval at `835` contains the maximum amount of steps taken during that 5 min.
```{r}
step_per_interval <- data_cleaned %>%
  group_by(interval) %>%
  summarize(avg=mean(steps))
plot.ts(step_per_interval$avg,ylab="Average Number of Steps Taken")  
step_per_interval$interval[step_per_interval$avg==max(step_per_interval$avg)]
```


## Imputing missing values
To impute missing values, I replace NAs with the average amount of steps taken during that time interval across all observations. The new mean remains the same, and the new median changes to ` 10766.19`, the same as the mean. There aren't significant changes in the histogram, although the new one has more data distributed at the center.
```{r}
nas <- filter(data,is.na(steps))
dim(nas)[1]
rest <- filter(data,!is.na(steps))
imputed <- merge (nas,step_per_interval)
replaced <- data.frame(steps=imputed$avg,date=as.factor(imputed$date),interval=imputed$interval)
new_data <- rbind(rest,replaced)
step_per_day_new <- new_data %>%
  group_by(date) %>%
  summarize(total=sum(steps))
hist(step_per_day_new$total)
mean(step_per_day_new$total)
median(step_per_day_new$total)
```

## Are there differences in activity patterns between weekdays and weekends?
As shown in the plot below, during weekdays there are more steps taken earlier in the day, whereas during weekends they are more evenly distributed across the day.
```{r}
new_data$date <- as.Date(new_data$date)
new_data$weekday <- weekdays(new_data$date)
new_data$if_weekday <- "weekday"
new_data$if_weekday[new_data$weekday=='Saturday']="weekend"
new_data$if_weekday[new_data$weekday=='Sunday']="weekend"
new_data$if_weekday <- factor(new_data$if_weekday)
step_per_interval_week <- new_data %>%
  group_by(if_weekday,interval) %>%
  summarize(avg=mean(steps))
#sub1 <- filter(step_per_interval_week,if_weekday=='weekday')
#sub2 <- filter(step_per_interval_week,if_weekday=='weekend')
#plot.ts(sub1$avg,ylab="Average Number of Steps Taken")  
#plot.ts(sub2$avg,ylab="Average Number of Steps Taken")  
```

```{r}
library(ggplot2)
g<- ggplot(step_per_interval_week, aes(x =interval , y=avg, color=if_weekday)) +
       geom_line() +
       labs(title = "Average Steps Per Day", x = "Interval", y = "Avg Steps") +
       facet_wrap(~if_weekday, ncol = 1, nrow=2)
g
```

